Windowed Developer Topic Analysis

### 2009-04-03 lda-paper.tex
abstract edit

  At different times, developers focus on different topics and tasks
  during development.  These different topics and tasks might
  be obvious to developers familar with the project, but other
  stakeholders, like managers, might be less aware of the topics of
  development. Imagine the scenario where developers unexpectedly had to deal with
  issues orthogonal to the features suggested, at the end of the
  iteration a manager would wonder where all the effort went to.
  Currently it is difficult for stakeholders to
  audit what actually occurred and when certain features or topics
  were focussed on. We propose a visualization of the
  changing development topics over time. Our approach uses a topic
  analysis tool like Latent Dirichlet Allocation (LDA) or Latent
  Semantic Indexing (LSI). Instead of applying the tool to the
  entire corpus we apply it to windows (periods) of the corpus. This allows us to
  see the evolving stream of topics of development occuring
  over time.

@@

Topic analysis, with respect to Software Control Systems (SCSs) like
CVS, Subversion, Git etc., is potentially valuable as it has many uses
for various stakeholders and uses.
Imagine a
development team has agreed to implement some features at the start of
an iteration. By the end of the iteration only some of the agreed upon
features are implemented. The developers swear they were focusing
solely on the features agreed upon. Topic analysis could help by
teasing out the distinct topics or focusses of the developers. Perhaps
the developers had to deal with issues.
Developers might have grouped those orthogonal changes into
one of the features they were working on, thus when asked about how
their time was spent the orthogonal changes would not be covered. This
is what we hope that windowed topic analysis can uncover. If windowed
topic analysis can discover important but overlooked topics, it can
help both developers and managers. Developer can see a
summary of what worked on, and managers can development.

Topic analysis of revisions in SCSs is important as often revisions
belonging to different topics are interleaved or even mixed inside of
commits.  Sometimes a single revision to a file contains multiple bug
fixes.  Using powerful topic analysis techniques like Latent Semantic
Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can seperate
these mixed comments into topics. These topics describe seperate ideas
that are interleaved throughtout the development. They indicate groups
of comments which are related to each other by their language, often
the content of the comments are related.  In this case a topic is a
set of tokens that are related to each other.

@@

Some terms we will be using for the rest of the paper include:
message, word distribution, topic and trend. \emph{Messages} are
blocks of text written by developers. In this paper messages will be
the CVS commit comments made when the user commits changes to files in
a repository. Word distributions are the summary of a message by word
count. The word distribution will be over all messages
looked at but since most words will not appear in a message, a word
distribution is effectively a word count divided by the message
size. A topic is a unique word distribution, but effectively a set of
words unique to an independant behaviour in the text corpus, in this
paper we often summarize topics by their top 10 more frequent words.
A trend is a one or more similar topics which reoccur over
time. Trends are important because they are sets of topics which
reoccur over the development of a project.

@@

Latent Dirichlet Alloation (LDA)~\cite{944937} is a competitor to
Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
(pLSI) and semantic clustering. It is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is model a documents as mixture model of topics.  The common use of
LDA in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

Each of these, LDA and LSI, extract topic clusters
from the documents that are independent from one another. LDA doesn't
name these clusters but the clusters consist of words whose
relationship is often obvious to someone familar with the
corpus. Technically we could swap LDA for LSI. Our data is posed as
documents with word distributions (word counts per documents) and LDA
extracts distinct topics (clusters of words) from the documents.

@@

LSI is related to LDA and has been used to identify topics in software
artifacts for forumal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
This is when concepts are automatically extracted from source code or
documents are related to part of the source code.  Concept location is
how high level concepts relate to low level entities like source code,
for instance when fixing a bug how do the concepts in the bug report
map back to the source code?  Semantic Clustering has also been used
for similar reason~\cite{1698774,1566153}.

Grant et al.~\cite{scottcordy} have used an alternative technique to
LSI and LDA, they used Independant Component Analysis to seperate
topic signals from software code.

@@

The first visualization is a compressed view, we try to sink the
larger trends to the bottom, once the larger trends are stacked, we
fill in the gaps with the smaller trends in the same window and then
stack the smaller trends on top. Although there is a chance that gaps
will be left due to nonoptimal stacking in practice there are lot of
small trends (90\% of all trends contain 1 topic) which fill in these
gaps quickly. The trend segments that share the same color are the
same trend, colors are randomly chosen for each trend, so color
similarity is meaningless, only exact color matching matters.
The compressed view is nice because it shows all of the
information and the repeating continuous trends are easy to pick out,
unfortunately repeating discontinuous trends are harder to spot.

The second visualization is the trend-time plot, it displays repeating
trends more clearly by dedicating a horizonal line for trend segments
dedicated to one trend. This means that if a trend contains
discontinuous segments the segments appear on the same line. One down
side to this view is the least common trends need to be pruned
otherwise it'd take too many horizontal lines to plot them.

@@

Once stop words were removed the some of hte purposes became more
clear. The mega trend disappeared, the largest trend became bug fixes
submitted by Dal Zotto which is a short trend.

@@

The plots we produced of MaxDB 7.500 were interesting, there was a
period where no real development occured and thus there were no topics or trends
whatsoever. We have two periods to evaluate for MaxDB 7.500, the first
period from June 2004 to jan 2005, and the second period from June
2005 to June 2006.

>>>>

%   At different times, developers focus on different topics and tasks
%   during development.  These different topics and tasks might
%   be obvious to developers familar with the project, but other
%   stakeholders, like managers, might be less aware of the topics of
%   development. Imagine the scenario where developers unexpectedly had to deal with
%   issues orthogonal to the features suggested, at the end of the
%   iteration a manager would wonder where all the effort went to.
%   Currently it is difficult for stakeholders to
%   audit what actually occurred and when certain features or topics
%   were focussed on. We propose a visualization of the
%   changing development topics over time. Our approach uses a topic
%   analysis tool like Latent Dirichlet Allocation (LDA) or Latent
%   Semantic Indexing (LSI). Instead of applying the tool to the
%   entire corpus we apply it to windows (periods) of the corpus. This allows us to
%   see the evolving stream of topics of development occuring
%   over time.

  During different periods, developers focus on different topics and
  different tasks during development. Stake-holders, such as managers,
  not directly involved with the development might have less awareness
  of the project and the topics and tasks that the developers are
  focussed on.  Imagine the scenario where a manager is trying to
  discover where all the developers' effort went after features have
  not been implemented because the developers unexpectedly had to deal
  with issues orthogonal to these features.  We propose to help
  stake-holder understand the topics of development by extracting
  topics from the descriptions of changes that programmers provide
  when they make changes.  We leverage topic analysis tools such as
  Latent Dirichlet Allocation (LDA) or Latent Semantic Indexing (LSI)
  to analyze periods, such as months, within a project's
  lifetime. Previous work applied LDA and LSI to
  to a project's entire lifetime.  These tools associate
  descriptions with independent topics extracted from these
  descriptions.  We propose a visualization of changing development
  topics over time that allows us to see the evolving stream of topics
  of development occurring over time. We also show that windowed topic
  analysis offers benefits compare to entire lifetime topic analysis
  because of the independence of topics between windows.

@@

Topic analysis, with respect to Software Control Systems (SCSs) like
CVS, Subversion, Git etc., is potentially valuable as it has many uses
for various stake-holders and uses.
Imagine a
development team has agreed to implement some features at the start of
an iteration. By the end of the iteration only some of the agreed upon
features are implemented. The developers swear they were focusing
solely on the features agreed upon. Topic analysis could help by
teasing out the distinct topics or focusses of the developers. Perhaps
the developers had to deal with issues.
Developers might have grouped those orthogonal changes into
one of the features they were working on, thus when asked about how
their time was spent the orthogonal changes would not be covered. This
is what we hope that windowed topic analysis can uncover. If windowed
topic analysis can discover important but overlooked topics, it can
help both developers and managers. Developer can see a
summary of what worked on, and managers can development.

Topic analysis of revisions in SCSs is important as often revisions
belonging to different topics are interleaved or even mixed inside of
commits.  Sometimes a single revision to a file contains multiple bug
fixes.  Using powerful topic analysis techniques like Latent Semantic
Indexing (LSI) or Latent Dirichlet Allocation (LDA) one can separate
these mixed comments into topics. These topics describe separate ideas
that are interleaved throughout the development. They indicate groups
of comments which are related to each other by their language, often
the content of the comments are related.  In this case a topic is a
set of tokens that are related to each other.

@@

Some terms we will be using for the rest of the paper include:
message, word distribution, topic and trend. \emph{Messages} are
blocks of text written by developers. In this paper messages will be
the CVS commit comments made when the user commits changes to files in
a repository. Word distributions are the summary of a message by word
count. The word distribution will be over all messages
looked at but since most words will not appear in a message, a word
distribution is effectively a word count divided by the message
size. A topic is a unique word distribution, but effectively a set of
words unique to an independent behavior in the text corpus, in this
paper we often summarize topics by their top 10 more frequent words.
A trend is a one or more similar topics which reoccur over
time. Trends are important because they are sets of topics which
reoccur over the development of a project.

@@

Latent Dirichlet Allocation (LDA)~\cite{944937} is a competitor to
Latent Semantic Indexing (LSI), probabilistic Latent Semantic Indexing
(pLSI) and semantic clustering. It is used for document modelling,
document clustering and collaborative filtering. What LDA attempts to
do is model a documents as mixture model of topics.  The common use of
LDA in software engineering
literature~\cite{lukins2008,10.1109/MSR.2007.20,NIPS2007637,1321709}
is to extract topic clusters from documents such as methods, bug
reports, source code and files.

Each of these, LDA and LSI, extract topic clusters
from the documents that are independent from one another. LDA doesn't
name these clusters but the clusters consist of words whose
relationship is often obvious to someone familiar with the
corpus. We could swap LDA for LSI as produce similar results. Our data is posed as
documents with word distributions (word counts per documents) and LDA
extracts distinct topics (clusters of words) from the documents.

@@

LSI is related to LDA and has been used to identify topics in software
artifacts for formal concept analysis and concept
location~\cite{1421013,1374321,10.1109/ICPC.2007.13,10.1109/ICPC.2006.17}.
This is when concepts are automatically extracted from source code or
documents are related to part of the source code.  Concept location is
how high level concepts relate to low level entities like source code,
for instance when fixing a bug how do the concepts in the bug report
map back to the source code?  Semantic Clustering has also been used
for similar reason~\cite{1698774,1566153}.

Grant et al.~\cite{scottcordy} have used an alternative technique to
LSI and LDA, they used Independent Component Analysis to separate
topic signals from software code.

@@

The first visualization is a compressed view, we try to sink the
larger trends to the bottom, once the larger trends are stacked, we
fill in the gaps with the smaller trends in the same window and then
stack the smaller trends on top. Although there is a chance that gaps
will be left due to non-optimal stacking in practice there are lot of
small trends (90\% of all trends contain 1 topic) which fill in these
gaps quickly. The trend segments that share the same color are the
same trend, colors are randomly chosen for each trend, so color
similarity is meaningless, only exact color matching matters.
The compressed view is nice because it shows all of the
information and the repeating continuous trends are easy to pick out,
unfortunately repeating discontinuous trends are harder to spot.

The second visualization is the trend-time plot, it displays repeating
trends more clearly by dedicating a horizontal line for trend segments
dedicated to one trend. This means that if a trend contains
discontinuous segments the segments appear on the same line. One down
side to this view is the least common trends need to be pruned
otherwise it'd take too many horizontal lines to plot them.

@@

Once stop words were removed the some of the purposes became more
clear. The mega- trend disappeared, the largest trend became bug fixes
submitted by Dal Zotto which is a short trend.

@@

The plots we produced of MaxDB 7.500 were interesting, there was a
period where no real development occurred and thus there were no topics or trends
whatsoever. We have two periods to evaluate for MaxDB 7.500, the first
period from June 2004 to jan 2005, and the second period from June
2005 to June 2006.
